---
title: "oPE-threshold-investigation"
output: html_notebook
---

```{r}

#----------------------------------------------------------------
#
# This code accompanies Gagl B., Sassenhagen J., Haan S., 
# Gregorova K., Richlan R. and Fiebach C.: Visual word 
# recognition relies on an orthographic prediction error signal
#
#----------------------------------------------------------------
#
#Estimation functions:
# prediction image
# orthographic prediction error: oPE
# number of pixel
#
#Evaluation datasets:
# Five letter words from German: "./ger5_fin.csv"
# Five letter words from English: "./us5_fin.csv"
# Five letter words from French: "./fr5_fin.csv"
# Five letter words from Dutch: "./du5_fin.csv"           
#
# Lexical decision data German #1: "./ger_tr_e1_v01.csv"
# Lexical decision data German #2; Noise experiment: "./ger_e2_deg_dat.csv"
# Lexical decision data British lexicon project: "./blp_tr_data_v01.csv"
# Lexical decision data French lexicon project: "./flp_tr_data_v01.csv"
#
# fMRI data from inferior frontal activation cluster: "./ifg_beta_activation.txt"
# fMRI data from medial frontal activation cluster: "./mf_beta_activation.txt"
#
# EEG data at 230 ms: "./EEG_230.csv"
# EEG data at 340 ms: "./EEG_340.csv"
# EEG data at 430 ms: "./EEG_430.csv"
#
#----------------------------------------------------------------
# Copyright by Benjamin Gagl, 2018
# This code is made publicly available under  CC-BY-NC-ND 
#----------------------------------------------------------------
#
# libraries needed for the script
# please install in case they are not availiable.
# use: install.package()

library(lme4) #Version: 1.1-15
library(ggplot2) #Version: 2.2.1
```

```{r}
#Lexicon Evaluation datasets:
# Five letter words from German
dat_ger_5 = read.csv("./ger5_fin.csv",header = T)
paste("German: oPE by Npix correlation:",cor(dat_ger_5$oPE,dat_ger_5$Npix))
paste("German: oPE by OLD correlation:",cor(dat_ger_5$oPE,dat_ger_5$OLD20))
paste("German: oPE by Frequency correlation:",cor(dat_ger_5$oPE,dat_ger_5$freq))

# Five letter words from English: "./us5_fin.csv"
dat_eng_5 = read.csv("./us5_fin.csv",header = T)
paste("English: oPE by Npix correlation:",cor(dat_eng_5$oPE,dat_eng_5$Npix))
paste("English: oPE by OLD correlation:",cor(dat_eng_5$oPE,dat_eng_5$OLD20))
paste("English: oPE by Frequency correlation:",cor(dat_eng_5$oPE,dat_eng_5$Freq))

# Five letter words from French: "./fr5_fin.csv"
dat_fr_5 = read.csv("./fr5_fin.csv",header = T)
paste("French: oPE by Npix correlation:",cor(dat_fr_5$oPE,dat_fr_5$Npix))
paste("French: oPE by OLD correlation:",cor(dat_fr_5$oPE,dat_fr_5$OLD20))
paste("French: oPE by Frequency correlation:",cor(dat_fr_5$oPE,dat_fr_5$Freq))

# Five letter words from Dutch: "./du5_fin.csv" 
dat_du_5 = read.csv("./du5_fin.csv",header = T)
paste("Dutch: oPE by Npix correlation:",cor(dat_du_5$oPE,dat_du_5$Npix))
paste("Dutch: oPE by OLD correlation:",cor(dat_du_5$oPE,dat_du_5$OLD20))
paste("Dutch: oPE by Frequency correlation:",cor(dat_du_5$oPE,dat_du_5$Freq))

```
```{r}
#Lexial decision behaviour
dat_lexdec = read.csv("./lex_dec_thresholded_ope.csv",header = T)
summary(dat_lexdec)
```
```{r}
#correlation between original oPe and different thresholds
corr_df = data.frame(dat_lexdec$ope, dat_lexdec$py_ope_ger_binary_th_10p, dat_lexdec$py_ope_ger_binary_th_30p, dat_lexdec$py_ope_ger_binary_th_50p, dat_lexdec$py_ope_ger_binary_th_70p, dat_lexdec$py_ope_ger_binary_th_90p)
cor(corr_df)
```
```{r}
#linear mixed model for reaction times using the 5-letter prediction with different thresholds
m_group_log = (lmer(log(rt)~scale(ope)*group_num+scale(npix)*group_num+err+freq+(1|vp)+(scale(ope)|string),data=dat_lexdec))
summary(m_group_log)
#10p
m_group_log_10p = (lmer(log(rt)~scale(py_ope_ger_binary_th_10p)*group_num+scale(npix)*group_num+err+freq+(1|vp)+(scale(py_ope_ger_binary_th_10p)|string),data=dat_lexdec))
summary(m_group_log_10p)
#30p
m_group_log_30p = (lmer(log(rt)~scale(py_ope_ger_binary_th_30p)*group_num+scale(npix)*group_num+err+freq+(1|vp)+(scale(py_ope_ger_binary_th_30p)|string),data=dat_lexdec))
summary(m_group_log_30p)
#50p
m_group_log_50p = (lmer(log(rt)~scale(py_ope_ger_binary_th_50p)*group_num+scale(npix)*group_num+err+freq+(1|vp)+(scale(py_ope_ger_binary_th_50p)|string),data=dat_lexdec))
summary(m_group_log_50p)
#70p
m_group_log_70p = (lmer(log(rt)~scale(py_ope_ger_binary_th_70p)*group_num+scale(npix)*group_num+err+freq+(1|vp)+(scale(py_ope_ger_binary_th_70p)|string),data=dat_lexdec))
summary(m_group_log_70p)
#90p
m_group_log_90p = (lmer(log(rt)~scale(py_ope_ger_binary_th_90p)*group_num+scale(npix)*group_num+err+freq+(1|vp)+(scale(py_ope_ger_binary_th_90p)|string),data=dat_lexdec))
summary(m_group_log_90p)

#Model comparison: 
aic_ger = anova(Null_model,m_group_log,m_group_log_10p,m_group_log_30p,m_group_log_50p,m_group_log_70p,m_group_log_90p)
aic_ger
```

```{r}
#linear mixed model for reaction times using the 5-letter prediction
m_group_log = (lmer(log(rt)~scale(ope)*group_num+scale(npix)*group_num+err+freq+(1|vp)+(scale(ope)|string),data=dat_lexdec))
summary(m_group_log)

#Model without the number of pixel parameters
m_group_log_only_oPE = (lmer(log(rt)~scale(ope)*group_num+err+freq+(1|vp)+(scale(ope)|string),data=dat_lexdec))
summary(m_group_log_only_oPE)

#Model without the oPE parameters
m_group_log_only_nPix = (lmer(log(rt)~scale(npix)*group_num+err+freq+(1|vp)+(scale(ope)|string),data=dat_lexdec))
summary(m_group_log_only_nPix)

#Null model without both parameters
Null_model = (lmer(log(rt)~group_num+err+freq+(1|vp)+(scale(ope)|string),data=dat_lexdec))
summary(Null_model)

#Model comparison: 
aic_ger = anova(Null_model,m_group_log_only_nPix,m_group_log_only_oPE,m_group_log)
```

```{r}
#Bar plot for model comparison

aic_diff = c(aic_ger$AIC[1]- aic_ger$AIC[2:length(aic_ger$AIC)])
Model = as.factor(c("10p","oPE only","Both"))
model_comp_plot_ger = data.frame(aic_diff,Model)

model_comp_plot_ger$Model = factor(model_comp_plot_ger$Model, levels=levels(model_comp_plot_ger$Model)[c(2,3,1)], ordered=TRUE)


r <- .5
t <- seq(0, 180, by = 1) * pi / 180
x <- r * cos(t)
y <- r*4 * sin(t)
y[20:162] <- y[20] # Flattens the arc

arc.df <- data.frame(Group = x, Value = y)

r <- .5
t <- seq(0, 180, by = 1) * pi / 180
x <- r * cos(t)
y <- r*4 * sin(t)
y[20:181] <- y[20] # Flattens the arc

arc.df_lng <- data.frame(Group = x, Value = y)
r <- .5
t <- seq(0, 180, by = 1) * pi / 180
x <- r * cos(t)
y <- r*4 * sin(t)
y[0:162] <- y[20] # Flattens the arc

arc.df_ri <- data.frame(Group = x, Value = y)


aic_pl_ger = ggplot(model_comp_plot_ger,aes(y=aic_diff,x=Model))+
  geom_bar(stat="identity")+
  theme_minimal()+
  ylab("AIC difference to H0")+
  xlab("Model")+
  geom_line(data = arc.df, aes(Group+1.5, Value+41), lty = 1) +
  geom_text(x = 1.5, y = 42, label = "*", cex=7)+
  geom_line(data = arc.df_lng, aes(Group+2.5, Value+45), lty = 1) +
  geom_line(data = arc.df_ri, aes(Group+1.5, Value+45), lty = 1) +
  geom_text(x = 2, y = 46, label = "*", cex=7)+
  ylim(0,46)

```
```

```{r}
#Code for Fig. 3a
#Estimate the oPE/Npixel effects without covariates
dat_remef = (na.omit(dat_lexdec))
dat_remef$ope = round(dat_remef$ope/10,0)*10
dat_remef$npix = round(dat_remef$npix/10,0)*10
dat_remef$category = factor(dat_remef$category,levels(dat_remef$category)[c(1,3,2)])

m_group = (lmer(log(rt)~scale(ope)*group_num+scale(npix)*group_num+err+(1|vp)+(scale(ope)|string),data=dat_lexdec))
#summary(m_group)

dat_remef$rt_remef = (keepef(m_group, fix = c("scale(ope)","group_num","scale(ope):group_num")))

#oPE figure
ope_group_int = ggplot(aggregate(rt_remef~ope+category,FUN=mean,dat_remef),aes(y=rt_remef,x=ope,color=category))+
  geom_point(alpha=.5)+
  geom_smooth(method=lm)+
  theme_minimal()+
  xlab('oPE')+
  ylab('Response time [log. ms]')+
  theme(legend.position = "None")+
  ylim(6.35,6.68)


#Npixel figure
dat_remef$rt_remef = (keepef(m_group, fix = c("scale(npix)","group_num","group_num:scale(npix)")))

npix_group_int = ggplot(aggregate(rt_remef~npix+category,FUN=mean,dat_remef),aes(y=rt_remef,x=npix,color=category))+
  geom_point(alpha=.5)+
  geom_smooth(method=lm)+
  theme_minimal()+
  xlab('Npixel')+
  ylab('Response time [log. ms]')+
  theme(legend.position = "None")+
  ylim(6.35,6.68)

grid.arrange(ope_group_int,npix_group_int,aic_pl_ger,ncol=3)
```

```{r}
#Linear mixed model including the Orthographic Levenshtein Distance (OLD)
m_group_log_old = (lmer(log(rt)~scale(ope)*group_num+scale(npix)*group_num+err+freq+old+(1|vp)+(scale(ope)|string),data=dat_lexdec))
summary(m_group_log_old)

#Linear mixed model including the oPE estimate based on the prediction including all word lengths (see Fig. 1e)
m_group_log_all = (lmer(log(rt)~scale(ope_all)*group_num+scale(npix)*group_num+err+freq+(1|vp)+(scale(ope_all)|string),data=dat_lexdec))
summary(m_group_log_all)
```



```{r}
#Lexial decision behaviour: English data
dat_lexdec_br = read.csv("./blp_tr_data_v01.csv",header = T)
summary(dat_lexdec_br)
#linear mixed model for reaction times using the 5-letter prediction
m_group_log_br = lmer(log(rt)~lex*scale(Npixel)+lex*scale(oPE)+freq+(1|participant)+(1|trial),data=dat_lexdec_br)
summary(m_group_log_br)

#model comparison: model without the number of pixel parameters
m_group_log_only_oPE_br = lmer(log(rt)~lex*scale(oPE)+freq+(1|participant)+(1|trial),data=dat_lexdec_br)
summary(m_group_log_only_oPE_br)

#model comparison: model without the number of oPE parameters
m_group_log_only_nPix_br = lmer(log(rt)~lex*scale(Npixel)+freq+(1|participant)+(1|trial),data=dat_lexdec_br)
summary(m_group_log_only_nPix_br)

#model comparison: Null model
m_group_log_NULL_br = lmer(log(rt)~lex+freq+(1|participant)+(1|trial),data=dat_lexdec_br)
summary(m_group_log_NULL_br)

aic_br = anova(m_group_log_NULL_br,m_group_log_only_nPix_br,m_group_log_only_oPE_br,m_group_log_br)
```

```{r}
#Bar plot for model comparison

aic_diff = c(aic_br$AIC[1]- aic_br$AIC[2:length(aic_br$AIC)])
Model = as.factor(c("Npix only","oPE only","Both"))
model_comp_plot_br = data.frame(aic_diff,Model)

model_comp_plot_br$Model = factor(model_comp_plot_br$Model, levels=levels(model_comp_plot_br$Model)[c(2,3,1)], ordered=TRUE)


aic_pl_br = ggplot(model_comp_plot_br,aes(y=aic_diff,x=Model))+
  geom_bar(stat="identity")+
  theme_minimal()+
  ylab("AIC difference to H0")+
  xlab("Model")+
  ylim(0,80)


```

```{r}
#Code for Fig. 3b
#Estimate the oPE/Npixel effects without covariates

dat_remef = (na.omit(dat_lexdec_br))
dat_remef$ope = round(dat_remef$oPE/10,0)*10
dat_remef$npix = round(dat_remef$Npixel/10,0)*10
dat_remef$category = as.factor(dat_remef$lex*-1)

m_group = lmer(log(rt)~lex*scale(Npixel)+lex*scale(oPE)+freq+(1|participant)+(1|trial),data=dat_remef)
#summary(m_group)

dat_remef$rt_remef = (keepef(m_group, fix = c("scale(oPE)","lex","lex:scale(oPE)")))

#oPE figure
ope_group_int_br = ggplot(aggregate(rt_remef~ope+category,FUN=mean,dat_remef),aes(y=rt_remef,x=ope,color=category))+
  geom_point(alpha=.5)+
  geom_smooth(method=lm)+
  theme_minimal()+
  xlab('oPE')+
  ylab('Response time [log. ms]')+
  theme(legend.position = "None")+
  ylim(6.28,6.58)+
  scale_color_manual(values=c("#53B64C","#6C9DF8"))


#Npixel figure
dat_remef$rt_remef = (keepef(m_group, fix = c("scale(Npixel)","lex","lex:scale(Npixel)")))

npix_group_int_br = ggplot(aggregate(rt_remef~npix+category,FUN=mean,dat_remef),aes(y=rt_remef,x=npix,color=category))+
  geom_point(alpha=.5)+
  geom_smooth(method=lm)+
  theme_minimal()+
  xlab('Npixel')+
  ylab('Response time [log. ms]')+
  theme(legend.position = "None")+
  ylim(6.28,6.58)+
  scale_color_manual(values=c("#53B64C","#6C9DF8"))

grid.arrange(ope_group_int_br,npix_group_int_br,aic_pl_br,ncol=3)
```

```{r}
#Lexial decision behaviour: French data
dat_lexdec_fr = read.csv("./flp_tr_data_v01.csv",header = T)
summary(dat_lexdec_fr)
```

```{r}
#linear mixed model for reaction times using the 5-letter prediction
m_group_log_fr_lex = lmer(log(rt)~lex_n*scale(Npixel)+lex_n*scale(ope)+freq_log+(1|vp)+(1|mot),data=dat_lexdec_fr)

summary(m_group_log_fr_lex)

#model comparison: model without the number of pixel parameters
m_group_log_only_oPE_fr = lmer(log(rt)~lex_n*scale(ope)+freq_log+(1|vp)+(1|mot),data=dat_lexdec_fr)

summary(m_group_log_only_oPE_fr)

#model comparison: model without the number of oPE parameters
m_group_log_only_nPix_fr = lmer(log(rt)~lex_n*scale(Npixel)+freq_log+(1|vp)+(1|mot),data=dat_lexdec_fr)

summary(m_group_log_only_nPix_fr)

#model comparison: model without the number of oPE parameters
m_group_log_NULL_fr = lmer(log(rt)~lex_n+freq_log+(1|vp)+(1|mot),data=dat_lexdec_fr)

summary(m_group_log_NULL_fr)


aic_fr = anova(m_group_log_NULL_fr,m_group_log_only_nPix_fr,m_group_log_only_oPE_fr,m_group_log_fr_lex)
```

```{r}
#Bar plot for model comparison

aic_diff = c(aic_fr$AIC[1]- aic_fr$AIC[2:length(aic_fr$AIC)])
Model = as.factor(c("Npix only","oPE only","Both"))
model_comp_plot_fr = data.frame(aic_diff,Model)

model_comp_plot_fr$Model = factor(model_comp_plot_fr$Model, levels=levels(model_comp_plot_fr$Model)[c(2,3,1)], ordered=TRUE)


aic_pl_fr = ggplot(model_comp_plot_fr,aes(y=aic_diff,x=Model))+
  geom_bar(stat="identity")+
  theme_minimal()+
  ylab("AIC difference to H0")+
  xlab("Model")+
  ylim(0,12)

```

```{r}
#code for Fig. 3c 
#estimate the ope effect without covariates
dat_remef = (na.omit(dat_lexdec_fr))
dat_remef$ope = round(dat_remef$ope/10,0)*10
dat_remef$npix = round(dat_remef$Npixel/10,0)*10
dat_remef$category = as.factor(dat_remef$lex_n*-1)

m_group = lmer(log(rt)~lex_n*scale(Npixel)+lex_n*scale(ope)+freq_log+(1|vp)+(1|mot),data=dat_remef)
#summary(m_group)

dat_remef$rt_remef = (keepef(m_group, fix = c("scale(ope)","lex_n","lex_n:scale(ope)")))

#oPE figure
ope_group_int_fr = ggplot(aggregate(rt_remef~ope+category,FUN=mean,dat_remef),aes(y=rt_remef,x=ope,color=category))+
  geom_point(alpha=.5)+
  geom_smooth(method=lm)+
  theme_minimal()+
  xlab('oPE')+
  ylab('Response time [log. ms]')+
  theme(legend.position = "None")+
  ylim(6.575,6.69)+
  scale_color_manual(values=c("#53B64C","#6C9DF8"))

 
#Npixel figure
dat_remef$rt_remef = (keepef(m_group, fix = c("scale(Npixel)","lex_n","lex_n:scale(Npixel)")))

npix_group_int_fr = ggplot(aggregate(rt_remef~npix+category,FUN=mean,dat_remef),aes(y=rt_remef,x=npix,color=category))+
  geom_point(alpha=.5)+
  geom_smooth(method=lm)+
  theme_minimal()+
  xlab('Npixel')+
  ylab('Response time [log. ms]')+
  theme(legend.position = "None")+
  ylim(6.575,6.69)+
  scale_color_manual(values=c("#53B64C","#6C9DF8"))


grid.arrange(ope_group_int_fr,npix_group_int_fr,aic_pl_fr,ncol=3)
```


```{r}
#Lexial decision behaviour: replication German
dat_lexdec_ger2 = read.csv("./ger_e2_deg_dat.csv",header = T)
summary(dat_lexdec_ger2)
```

```{r}
#Model estimation for presentaion without noise
#oPE only model
ope_only_deg_lum_ope_model_deg0 = lmer(log(time)~scale(ope)*group_num+freq+err+(1|vp)+(1|stim),subset(dat_lexdec_ger2,dat_lexdec_ger2$cond=="0%"))

#Npixel only model
npix_only_deg_lum_ope_model_deg0 = (lmer(log(time)~scale(npix)*group_num+freq+err+(1|vp)+(1|stim),subset(dat_lexdec_ger2,dat_lexdec_ger2$cond=="0%")))

#Full model
deg_lum_ope_model_deg0 = (lmer(log(time)~scale(ope)*group_num+scale(npix)*group_num+freq+err+(1|vp)+(1|stim),subset(dat_lexdec_ger2,dat_lexdec_ger2$cond=="0%")))

summary(deg_lum_ope_model_deg0)

#Null model
NULL_deg_lum_ope_model_deg0 = (lmer(log(time)~group_num+freq+err+(1|vp)+(1|stim),subset(dat_lexdec_ger2,dat_lexdec_ger2$cond=="0%")))

aic_deg0 = anova(NULL_deg_lum_ope_model_deg0,npix_only_deg_lum_ope_model_deg0,ope_only_deg_lum_ope_model_deg0,deg_lum_ope_model_deg0)
```



```{r}
#Model estimation for presentaion low noise levels
#oPE only model
ope_only_deg_lum_ope_model_deg20 = lmer(log(time)~scale(ope)*group_num+freq+err+(1|vp)+(1|stim),subset(dat_lexdec_ger2,dat_lexdec_ger2$cond=="20%"))

#Npixel only model
npix_only_deg_lum_ope_model_deg20 = (lmer(log(time)~scale(npix)*group_num+freq+err+(1|vp)+(1|stim),subset(dat_lexdec_ger2,dat_lexdec_ger2$cond=="20%")))

#Full model
deg_lum_ope_model_deg20 = (lmer(log(time)~scale(ope)*group_num+scale(npix)*group_num+freq+err+(1|vp)+(1|stim),subset(dat_lexdec_ger2,dat_lexdec_ger2$cond=="20%")))

summary(deg_lum_ope_model_deg20)

#Null model
NULL_deg_lum_ope_model_deg20 = (lmer(log(time)~group_num+freq+err+(1|vp)+(1|stim),subset(dat_lexdec_ger2,dat_lexdec_ger2$cond=="20%")))

aic_deg20 = anova(NULL_deg_lum_ope_model_deg20,npix_only_deg_lum_ope_model_deg20,ope_only_deg_lum_ope_model_deg20,deg_lum_ope_model_deg20)
```

```{r}
#Model estimation for presentaion high noise levels
#oPE only model
ope_only_deg_lum_ope_model_deg40 = lmer(log(time)~scale(ope)*group_num+freq+err+(1|vp)+(1|stim),subset(dat_lexdec_ger2,dat_lexdec_ger2$cond=="40%"))

#Npixel only model
npix_only_deg_lum_ope_model_deg40 = (lmer(log(time)~scale(npix)*group_num+freq+err+(1|vp)+(1|stim),subset(dat_lexdec_ger2,dat_lexdec_ger2$cond=="40%")))

#Full model
deg_lum_ope_model_deg40 = (lmer(log(time)~scale(ope)*group_num+scale(npix)*group_num+freq+err+(1|vp)+(1|stim),subset(dat_lexdec_ger2,dat_lexdec_ger2$cond=="40%")))

summary(deg_lum_ope_model_deg40)

#Null model
NULL_deg_lum_ope_model_deg40 = (lmer(log(time)~group_num+freq+err+(1|vp)+(1|stim),subset(dat_lexdec_ger2,dat_lexdec_ger2$cond=="40%")))

aic_deg40 = anova(NULL_deg_lum_ope_model_deg40,npix_only_deg_lum_ope_model_deg40,ope_only_deg_lum_ope_model_deg40,deg_lum_ope_model_deg40)
```

```{r}
#Bar plot for model comparison

aic_diff = c(aic_deg0$AIC[1]- aic_deg0$AIC[2:length(aic_deg0$AIC)]
             ,aic_deg20$AIC[1]- aic_deg20$AIC[2:length(aic_deg20$AIC)]
             ,aic_deg40$AIC[1]- aic_deg40$AIC[2:length(aic_deg40$AIC)])
Model = as.factor(rep(c("Npix\nonly","oPE\nonly","Both"),3))
deg = as.factor(c(rep("0%",3),rep("20%",3),rep("40%",3)))
model_comp_plot = data.frame(aic_diff,Model,deg)

model_comp_plot$Model = factor(model_comp_plot$Model, levels=levels(model_comp_plot$Model)[c(2,3,1)], ordered=TRUE)

aic_pl_deg_bar = ggplot(model_comp_plot,aes(y=aic_diff,x=Model))+
  geom_bar(stat="identity")+
  theme_minimal()+
  ylab("AIC difference to H0")+
  xlab("")+
  facet_wrap(~deg,nrow=1)+
  ylim(-5,14)+
  ggtitle(label="Model\ncomparisons")

```

```{r}
#code for Fig. 4
#estimate the ope effect without covariates
dat_remef = (na.omit(dat_lexdec_ger2))
dat_remef$ope_ = round(dat_remef$ope/10,0)*10
dat_remef$npix_ = round(dat_remef$npix/10,0)*10
dat_remef$category = factor(dat_remef$lex_cond, levels=levels(dat_remef$lex_cond)[c(1,3,2)], ordered=TRUE)
dat_remef$cond_int = 0
dat_remef$cond_int[dat_remef$cond=="20%"] = 20
dat_remef$cond_int[dat_remef$cond=="40%"] = 40

m_group = (lmer(log(time)~scale(ope)*group_num*cond_int+scale(npix)*group_num*cond_int+freq+err+(1|vp)+(1|stim),dat_remef))
#summary(m_group)

dat_remef$rt_remef = (keepef(m_group
                             , fix = c("scale(ope)"
                                       ,"group_num"
                                       ,"cond_int"
                                       ,"scale(ope):group_num"
                                       ,"scale(ope):cond_int"
                                       ,"scale(ope):group_num:cond_int")))

#oPE figure
ag_plot = aggregate(rt_remef~ope_+category+cond,FUN=mean,dat_remef)
ope_group_int = ggplot(ag_plot,aes(y=rt_remef,x=ope,color=category))+
  geom_point(alpha=.5)+
  geom_smooth(method=lm)+
  theme_minimal()+
  xlab('oPE')+
  ylab('Response time [log. ms]')+
  theme(legend.position = "None")+
  facet_wrap(~cond,ncol = 1)#+
  #ylim(400,1200)

ope_group_int_deg0 = ggplot(subset(ag_plot,ag_plot$cond=="0%"),aes(y=rt_remef,x=ope_,color=category))+
  geom_point(alpha=.5)+
  geom_smooth(method=lm)+
  theme_minimal()+
  xlab('oPE')+
  ylab('Response time [log. ms]')+
  theme(legend.position = "None")+
  ylim(6.1,7.2)+
  ggtitle(label = "oPE effect\n0% degradiation")


#Npixel figure
dat_remef$rt_remef = (keepef(m_group
                             , fix = c("scale(npix)"
                                       ,"group_num"
                                       ,"cond_int"
                                       ,"group_num:scale(npix)"
                                       ,"cond_int:scale(npix)"
                                       ,"group_num:cond_int:scale(npix)")))

ag_plot = aggregate(rt_remef~npix_+category+cond,FUN=mean,dat_remef)
npix_group_int = ggplot(ag_plot,aes(y=rt_remef,x=npix,color=category))+
  geom_point(alpha=.5)+
  geom_smooth(method=lm)+
  theme_minimal()+
  xlab('Npixel')+
  ylab('Response time [log. ms]')+
  theme(legend.position = "None")+
  facet_wrap(~cond,ncol = 1)#+
  #ylim(400,1200)

npix_group_int_deg40 = ggplot(subset(ag_plot,ag_plot$cond=="40%"),aes(y=rt_remef,x=npix_,color=category))+
  geom_point(alpha=.5)+
  geom_smooth(method=lm)+
  theme_minimal()+
  xlab('Npixel')+
  ylab('Response time [log. ms]')+
  theme(legend.position = "None")+
  ylim(6.1,7.2)+
  ggtitle(label = "Npixel effect\n40% degradiation")

grid.arrange(ope_group_int_deg0,npix_group_int_deg40,aic_pl_deg_bar,widths=c(1,1,2))
```

```{r}
#fMRI data

ifg = read.table("ifg_beta_activation.txt",header=T)
mf = read.table("mf_beta_activation.txt",header=T)

ifg$Lexicality = as.factor(c(rep("PW",39),rep("W",39)))
mf$Lexicality = as.factor(c(rep("PW",39),rep("W",39)))

ifg$vp = c(seq(1,39,1),seq(1,39,1))
mf$vp = c(seq(1,39,1),seq(1,39,1))

mf$Lexicality = factor(mf$Lexicality,levels = c("W","PW"))
ifg$Lexicality = factor(ifg$Lexicality,levels = c("W","PW"))


```

```{r}
#IFG ROI Figure
empty <- ggplot()+geom_point(aes(1,1), colour="white") +
     theme(                              
       plot.background = element_blank(), 
       panel.grid.major = element_blank(), 
       panel.grid.minor = element_blank(), 
       panel.border = element_blank(), 
       panel.background = element_blank(),
       axis.title.x = element_blank(),
       axis.title.y = element_blank(),
       axis.text.x = element_blank(),
       axis.text.y = element_blank(),
       axis.ticks = element_blank()
     )


box_sca = ggplot(ifg,aes(y=beta,x=Lexicality,color=Lexicality,fill=Lexicality))+
  #geom_histogram()
  theme_minimal( base_family = "")+
  geom_boxplot(alpha=.3)+
  geom_point(alpha=.3)+
  geom_line(alpha=.3,aes(group=vp),color=gray(.5))+
  xlab('')+
  ylab('oPE BOLD change [Beta weights]')+
  scale_color_manual(values = c("#4DAF4A","#377EB8"))+
  scale_fill_manual(values = c("#4DAF4A","#377EB8"))+
  theme(legend.position = 'none')



density = ggplot(ifg,aes(x=beta,color=Lexicality,fill=Lexicality))+
  #geom_histogram()+
  theme_minimal( base_family = "")+
  geom_density(alpha=.3)+
  xlab('')+
  ylab('Density')+
  scale_color_manual(values = c("#4DAF4A","#377EB8"))+
  scale_fill_manual(values = c("#4DAF4A","#377EB8"))+
  theme(legend.position = 'none')+ 
  coord_flip()

grid.arrange(empty,empty, box_sca, density, ncol=2, nrow=2, widths=c(4, 3), heights=c(1, 4))
```

```{r}
#MF Figure

empty <- ggplot()+geom_point(aes(1,1), colour="white") +
     theme(                              
       plot.background = element_blank(), 
       panel.grid.major = element_blank(), 
       panel.grid.minor = element_blank(), 
       panel.border = element_blank(), 
       panel.background = element_blank(),
       axis.title.x = element_blank(),
       axis.title.y = element_blank(),
       axis.text.x = element_blank(),
       axis.text.y = element_blank(),
       axis.ticks = element_blank()
     )


box_sca = ggplot(mf,aes(y=beta,x=Lexicality,color=Lexicality,fill=Lexicality))+
  #geom_histogram()
  theme_minimal( base_family = "")+
  geom_boxplot(alpha=.3)+
  geom_point(alpha=.3)+
  geom_line(alpha=.3,aes(group=vp),color=gray(.5))+
  xlab('')+
  ylab('oPE BOLD change [Beta weights]')+
  scale_color_manual(values = c("#4DAF4A","#377EB8"))+
  scale_fill_manual(values = c("#4DAF4A","#377EB8"))+
  theme(legend.position = 'none')



density = ggplot(mf,aes(x=beta,color=Lexicality,fill=Lexicality))+
  #geom_histogram()+
  theme_minimal( base_family = "")+
  geom_density(alpha=.3)+
  xlab('')+
  ylab('Density')+
  scale_color_manual(values = c("#4DAF4A","#377EB8"))+
  scale_fill_manual(values = c("#4DAF4A","#377EB8"))+
  theme(legend.position = 'none')+ 
  coord_flip()

grid.arrange(empty,empty, box_sca, density, ncol=2, nrow=2, widths=c(4, 3), heights=c(1, 4))
```

```{r}
#EEG: oPE effect at 230 ms at posterior sensors
dat_230 = read.csv("EEG_230.csv",header = T) #get dataset
dat_post = subset(dat_230,dat_230$cluster=='P') #select sensors

#Run linear mixed model
mod_n2_cn = lmer(mv~scale(ope)*scale(cond_num)+ scale(npix)*scale(cond_num)+(1|subject)+(1|string),data=dat_post)
summary(mod_n2_cn)
```

```{r}
#Focus on effect of interest by discarding confounding effects
dat_post$mv_ke = (keepef(mod_n2_cn, fix = c("scale(ope)","scale(cond_num)","scale(ope):scale(cond_num)")))

#Bin oPE
dat_post$ope_r = round(dat_post$ope/50,1)*50
dat_agg_pl = aggregate(mv_ke~ope_r+cond,dat_post,FUN=mean)


#Plot figue
ggplot(dat_agg_pl,aes(y=mv_ke,x=ope_r,color=cond))+
  geom_point(alpha=.1)+
  geom_smooth(method=lm)+
  theme_minimal()+
  ylab( expression(paste("Cluster pattern [",mu, "V]")))+
  xlab("oPE")+
  scale_color_manual(values = c("#F8766D","#619CFF","#00BA38"))
```

```{r}
#EEG: Word/Non-word effect at 340 ms at posterior sensors
dat_340 = read.csv("EEG_340.csv",header = T) #get dataset
dat_post_340 = subset(dat_340,dat_340$cluster=='P') #select sensors

#Run linear mixed model
mod_n3_cn = lmer(mv~scale(ope)*scale(cond_num)+ scale(npix)*scale(cond_num)+(1|subject)+(1|string),data=dat_post_340)
summary(mod_n3_cn)
```

```{r}
#Focus on effect of interest by discarding confounding effects
dat_post_340$mv_ke = (keepef(mod_n3_cn, fix = c("scale(cond_num)")))

#Aggregate data and rename factors
dat_agg_pl = aggregate(mv_ke~subject+cond+cond_num,dat_post_340,FUN=mean,na.rm=T)
dat_agg_pl$cond_abb = "CS"
dat_agg_pl$cond_abb[dat_agg_pl$cond_num==.5] = "PW"
dat_agg_pl$cond_abb[dat_agg_pl$cond_num==1] = "W"
dat_agg_pl$cond_abb = as.factor(dat_agg_pl$cond_abb)
dat_agg_pl$cond = as.character(dat_agg_pl$cond)
dat_agg_pl$cond[dat_agg_pl$cond_num==0] = "Consonant String"
dat_agg_pl$cond[dat_agg_pl$cond_num==0.5] = "Pseudoword"
dat_agg_pl$cond = as.factor(dat_agg_pl$cond)

#Plot figue
ggplot(dat_agg_pl,aes(y=mv_ke*10^6,x=cond_abb,color=cond,fill=cond))+
  geom_boxplot(alpha=.3)+
  geom_point(alpha=.3)+
  theme_minimal()+
  ylab( expression(paste("Cluster pattern [",mu, "V]")))+
  xlab("")+
  scale_color_manual(values = c("#F8766D","#619CFF","#00BA38"))+
  scale_fill_manual(values = c("#F8766D","#619CFF","#00BA38"))
```

```{r}
#EEG: oPE effect at 230 ms at posterior sensors
dat_430 = read.csv("EEG_430.csv",header = T) #get dataset
dat_front = subset(dat_430,dat_430$cluster=='F') #select sensors

#Run linear mixed model
mod_n4_cn = lmer(mv~scale(ope)*scale(cond_num)+ scale(npix)*scale(cond_num)+(1|subject)+(1|string),data=dat_front)
summary(mod_n4_cn)
```



```{r}
#Focus on effect of interest by discarding confounding effects
dat_front$mv_ke = (keepef(mod_n4_cn, fix = c("scale(ope)","scale(cond_num)","scale(ope):scale(cond_num)")))

#Bin oPE
dat_front$ope_r = round(dat_front$ope/50,1)*50
dat_agg_pl = aggregate(mv_ke~ope_r+cond,dat_front,FUN=mean)

#Plot figue
ggplot(dat_agg_pl,aes(y=mv_ke,x=ope_r,color=cond))+
  geom_point(alpha=.1)+
  geom_smooth(method=lm)+
  theme_minimal()+
  ylab(expression(paste("Cluster pattern [",mu, "V]")))+
  xlab("oPE")+
  scale_color_manual(values = c("#F8766D","#619CFF","#00BA38"))
```

```{r}
#Handwriting data
hw_ratings = read.csv("readability_rating.csv",header=T)
script_chars = read.csv("script_characteristics.csv",header=T)

#do the figures
m_pred_vpe_n = ggplot(script_chars,aes(x=1-meanpred,y=ope_n_script))+
  geom_point(aes(color=script))+
  geom_smooth(method=lm)+
  theme_minimal()+ 
  xlab('Prediction strength\n[mean gray value]')+
  ylab('oPE - Npixel normalized')+
  theme(legend.position = 'None')+ 
  scale_x_continuous( breaks=c(0.06,0.08,0.1) )

n_pred_vpe_n = ggplot(script_chars,aes(x=n_non1,y=ope_n_script))+
  geom_point(aes(color=script))+
  geom_smooth(method=lm)+
  theme_minimal()+ 
  xlab('Prediction precision\n[N gray pixels]')+
  ylab('oPE - Npixel normalized')+
  theme(legend.position = 'None')

readabilitypl_indi = ggplot(hw_ratings,aes(y=readability,x=ope_norm))+
  geom_point(alpha=.1)+
  geom_smooth(method = lm,alpha=0,aes(color=as.factor(vp)),size=.25)+
  geom_smooth(method = lm)+
  theme_minimal()+
  ylab("Script readability rating")+
  xlab("oPE - \nNpixel normalized")+
  theme(legend.position = 'None')+
  ylim(0.9,6.1)+ 
  scale_x_continuous( breaks=c(1.4,1.5,1.6) )

grid.arrange(m_pred_vpe_n,n_pred_vpe_n,readabilitypl_indi,ncol=3)
```

```{r}
#Handwriting analysis
summary(lmer(readability~(ope_norm)+((ope_norm)|vp),data = hw_ratings))
summary(lmer(ope_n_script~scale(n_non1)+scale(meanpred)+case+(1|script),data = script_chars))
```

